Microbiome analysis - Mirzapur follow-up

### 2020-09-08 code/R_markdown/P1_analysis_prep_figures.Rmd
added species random forest and updated figure 1

>>>>

Species
```{r}

#all
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F))$LIB13.C..hominis<40,1,0) %>% table()
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F))$C..mel.y<40,1,0) %>% table()
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F))$Pan_crypto<40,1,0) %>% table()

# diarrheal
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F & symptomatic == T))$LIB13.C..hominis<40,1,0) %>% table()
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F & symptomatic == T))$C..mel.y<40,1,0) %>% table()
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F & symptomatic == T))$Pan_crypto<40,1,0) %>% table()

# subclin
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F & symptomatic == F))$LIB13.C..hominis<40,1,0) %>% table()
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F & symptomatic == F))$C..mel.y<40,1,0) %>% table()
ifelse(sample_data(subset_samples(ps_rare_diff, pre_detection == F & symptomatic == F))$Pan_crypto<40,1,0) %>% table()

### 2020-09-08 code/R_markdown/P3_machine_learning.Rmd
added species random forest and updated figure 1

>>>>

## DATA NOT SHOWN: Does microbiome predict Crypto species at TOD? HERE

compare <- c(10,12) # TOD +/- symptoms
ps_for_class2 <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class2), as(sample_data(ps_for_class2),"data.frame") %>% select(Age.in.days))
y_hominis <- as.factor(ifelse(sample_data(ps_for_class2)$LIB13.C..hominis <40,1,0))
y_meleagridis <- as.factor(ifelse(sample_data(ps_for_class2)$C..mel.y <40,1,0))

rf = randomForest(x,y_hominis,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.0001,100))
rf
predictions = as.data.frame(predict(rf,type="prob"))
predictions$predict = names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
predictions$observed = y_hominis
ROC_res = roc(ifelse(predictions$observed==0, 0, 1), as.numeric(predictions$`0`))
auc(ROC_res)
save(list = c("rf","x","y_hominis"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_hominis_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

rf = randomForest(x,y_meleagridis,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.0001,100))
rf
predictions = as.data.frame(predict(rf,type="prob"))
predictions$predict = names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
predictions$observed = y_meleagridis
ROC_res = roc(ifelse(predictions$observed==0, 0, 1), as.numeric(predictions$`0`))
auc(ROC_res)
save(list = c("rf","x","y_meleagridis"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_meleagridis_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

### 2020-09-08 code/R_markdown/P5_mirzapur_analysis_figures.Rmd
added species random forest and updated figure 1

# load packages
library(ggplot2); library(phyloseq)

@@

>>>>

# load packages
library(ggplot2); library(phyloseq); library(dplyr)

@@

Species
```{r}

load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
crypto_cts = burden_all_path %>% filter(crypto == 1) %>%
  mutate(date_of_collection = as.Date(gsub('.{8}$', '', as.character(DOC)), "%m/%d/%Y")) %>%
  select(CHILDID, date_of_collection, CRYCTRT, type) %>% filter(CHILDID %in% burden_mirpur$SID)
rm(list= ls()[! (ls() %in% c('ps_mirzapur_rare','crypto_cts'))])

tempsampledata = sample_data(ps_mirzapur_rare)
tempsampledata = merge(tempsampledata, crypto_cts %>% select(-type), by.x = c("Child.ID","Date.of.collection"), by.y = c("CHILDID","date_of_collection"), all.x = T)
tempsampledata[tempsampledata$Clinical.type %in% c(1,2),"CRYCTRT"] = 45 # if predetection, Crypto DNA was not assessed or detected

# get other Ct values (targeted qPCRs)
all_ct = read.table("~/local_documents/work/crypto_microbiome/data_cannot_be_shared/Cq_values_all.txt",header=TRUE, sep='\t',comment.char = "",stringsAsFactors = F, na.strings = c("#N/A"))
all_ct[is.na(all_ct)] = 45 # replace #NA and real NAs with 0
all_ct[all_ct == 0] = 45 # replace #NA and real NAs with 0
tempsampledata = tempsampledata %>% mutate(sample_id_mod = substr(sample.id,1,nchar(sample.id)-5))
tempsampledata = merge(tempsampledata, all_ct, by.x = "sample_id_mod", by.y = "sample.id", all.x = T)
rownames(tempsampledata) = tempsampledata$use_as_rows; tempsampledata$use_as_rows = NULL

#12=diarrhea; 10=Clinic Visit; 1 =pre-clinic visit positive; 2=pre-diarrhea case positive

#all
ifelse(tempsampledata %>% filter(Clinical.type %in% c(10,12)) %>% select(LIB13.C..hominis)<40,1,0) %>% table()
ifelse(tempsampledata %>% filter(Clinical.type %in% c(10,12)) %>% select(C..mel.y)<40,1,0) %>% table()
ifelse(tempsampledata %>% filter(Clinical.type %in% c(10,12)) %>% select(Pan_crypto)<40,1,0) %>% table()

# diarrheal
ifelse(tempsampledata %>% filter(Clinical.type %in% c(12)) %>% select(LIB13.C..hominis)<40,1,0) %>% table()
ifelse(tempsampledata %>% filter(Clinical.type %in% c(12)) %>% select(C..mel.y)<40,1,0) %>% table()
ifelse(tempsampledata %>% filter(Clinical.type %in% c(12)) %>% select(Pan_crypto)<40,1,0) %>% table()

# subclin
ifelse(tempsampledata %>% filter(Clinical.type %in% c(10)) %>% select(LIB13.C..hominis)<40,1,0) %>% table()
ifelse(tempsampledata %>% filter(Clinical.type %in% c(10)) %>% select(C..mel.y)<40,1,0) %>% table()
ifelse(tempsampledata %>% filter(Clinical.type %in% c(10)) %>% select(Pan_crypto)<40,1,0) %>% table()
